Non-word repetition in Yélî Dnye

### 2021-09-15 manuscript.Rmd
end r3

As for birth order, much of the work on correlations between birth order and cognitive development (including language) has been carried out in the last 70 years and in agrarian or industrialized settings [@gratz2018competition;@barclay2015within], where nuclear families are more likely to be the prevalent rearing environment [@lancy2015anthropology]. It is possible that birth order differences are stronger in such a setting, because much of the stimulation can only come from the parents, and when there are multiple children, the inter-birth interval is small enough that older siblings may not be of an age that allows them to contribute to their younger siblings' stimulation. This contrasts with the picture just drawn in the Yélî community, where children will typically benefit from a rich and extensive socially stimulating setting, surrounded by siblings, and cousins of several orders, regardless of their birth order in their nuclear family.

<!-- ac2mc shouldn't we totally rewrite this? -->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is also relatively infrequent in this community [@casillas2020early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2020early]. Thus, if input quantities in early childhood are a major determinant of NWR scores, we should observe similarly low NWR scores as in @cristia2020infant.

@@

We addressed our first research question in a second way, by investigating patterns of error, looking at all attempts so as to base our generalizations on more data. There were no cases of insertion, and deletions were very rare: there were only `r sum(del_tab[c(oralVs,nasalVs)])` instances of deleted vowels (~`r mean(del_pc[c(oralVs,nasalVs)],na.rm=T)`% of all vowel targets), and `r sum(del_tab[c(simplexCs,complexCs)])` instances of deleted consonants (~`r mean(del_pc[c(simplexCs,complexCs)],na.rm=T)`% of all consonant targets). We therefore focus our qualitative description here on substitutions: There were `r dim(sub_bank)[1]` cases of substitutions, ~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1])*100,2)` of the `r dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1]` phones found collapsing across all children and target words, so that substitutions constituted the  majority of incorrect phones (~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] )*100,2)` of unmatched phones). To inform our understanding of how cross-linguistic patterns may be reflected in NWR scores, we asked: Is it the case that cross-linguistically less common and/or more complex phones are more frequently mispronounced, and more frequently substituted by more common ones than vice versa?

@@

Before closing, we would like to point out some salient limitations of the current work. To begin with, we only employed one set of non-words, in which not all characteristics that previous work suggest matter were manipulated [@chiat2015non]. As a result, we only have a rather whole-sale measure of performance, and we do not know to what extent lexical knowledge, pure phonological knowledge, and working memory, among others, contribute to children's performance. Similarly, our items varied systematically in length and typological frequency of the sounds included, but not in other potential dimensions (such as whether the items contained morphemes of the language or not). Additionally, we only had a single person interacting with children as well as interpreting children's production, so we do not know to what extent our findings generalize to other experimenters and research assistants. In addition, since both stimuli presentation and production data collected were audio-only, neither the children nor our research assistant were able to integrate visual cues in their interpretation. Although we know from other work that adults' perceptual performance on these types of sounds is well above chance from audio-only presentation [REF],  language processing for the majority of children will be audiovisual in natural conditions, and thus it may be interesting in the future to capture this aspect of speech.

>>>>

As for birth order, much of the work on correlations between birth order and cognitive development (including language) has been carried out in the last 70 years and in agrarian or industrialized settings [@gratz2018competition;@barclay2015within], where nuclear families were more likely to be the prevalent rearing environment [@lancy2015anthropology]. It is possible that birth order differences are stronger in such a setting, because much of the stimulation can only come from the parents. These effects may be much smaller in cultures where it is common for children to attend daycare at an early age (such as France) or where extended family typically live close by. The Yélî community falls in the latter type, as children are typically surrounded by siblings and cousins of several orders, regardless of their birth order in their nuclear family.

<!-- ac2mc shouldn't we totally rewrite this? // actually, this is all factually correct-->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is also relatively infrequent in this community [@casillas2020early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2020early]. Thus, if input quantities in early childhood are a major determinant of NWR scores, we should observe similarly low NWR scores as in @cristia2020infant.

@@

We addressed our first research question in a second way, by investigating patterns of error, looking at all attempts and not excluding errors resulting in real words, so as to base our generalizations on more data. Insertions and deletions were very rare (metathesis was not attested): there were only `r sum(del_tab[c(oralVs,nasalVs)])` instances of deleted vowels (~`r mean(del_pc[c(oralVs,nasalVs)],na.rm=T)`% of all vowel targets), and `r sum(del_tab[c(simplexCs,complexCs)])` instances of deleted consonants (~`r mean(del_pc[c(simplexCs,complexCs)],na.rm=T)`% of all consonant targets). We therefore focus our qualitative description here on substitutions: There were `r dim(sub_bank)[1]` cases of substitutions, ~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1])*100,2)` of the `r dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1]` phones found collapsing across all children and target words, so that substitutions constituted the  majority of incorrect phones (~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] )*100,2)` of unmatched phones). To inform our understanding of how cross-linguistic patterns may be reflected in NWR scores, we asked: Is it the case that cross-linguistically less common and/or more complex phones are more frequently mispronounced, and more frequently substituted by more common ones than vice versa?

@@

Before closing, we would like to point out some salient limitations of the current work. To begin with, we only employed one set of non-words, in which not all characteristics that previous work suggest matter were manipulated [@chiat2015non]. As a result, we only have a rather whole-sale measure of performance, and we do not know to what extent lexical knowledge, pure phonological knowledge, and working memory, among others, contribute to children's performance. Similarly, our items varied systematically in length and typological frequency of the sounds included, but not in other potential dimensions (such as whether the items contained morphemes of the language or not).

We relied on a single resource, PHOIBLE, for our estimation of typological frequency, and some readers may be worried about the effects of this choice. As far as we know, PHOIBLE is the most extensive archive of phonological inventories, so it is a reasonable choice in the current context. However, one may want to calculate typological frequency not by trying to have as many languages represented as possible, but rather by selecting a sample of typologically independent languages. In addition, it is not the case that all the world's languages are represented, and indeed some of the Yélî sounds were not found there. While these limitations should be borne in mind, it is important to also consider what our conclusions were, and that is that there is a non-trivial correlation between NWR and typological frequency. At present, we do not see how imbalance in the typological selection and missing data can conspire to produce the correlation we observe. If anything, these factors should increase  noise in the typological frequency estimation, in which case the correlation size we uncover is an underestimation of the true correlation.

Additionally, we only had a single person interacting with children as well as interpreting children's production, so we do not know to what extent our findings generalize to other experimenters and research assistants. Furthermore, since both stimuli presentation and production data collected were audio-only, neither the children nor our research assistant were able to integrate visual cues in their interpretation. Although we know from other work that adults' perceptual performance on these types of sounds is well above chance from audio-only presentation [REF],  language processing for the majority of children will be audiovisual in natural conditions, and thus it may be interesting in the future to capture this aspect of speech.
